---
title: "3. Data visualisation"
output: html_document
---

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
``` 

### 3.2.4 Exercises

1. Run `ggplot(data = mpg)`. What do you see?

```{r}
ggplot(data = mpg)
```

Nothing. No layers have been added to `ggplot()`.

2. How many rows are in `mpg`? How many columns?

232 rows and 11 columns in total.

```{r}
mpg
```

3. What does the `drv` variable describe? Read the help for `?mpg` to find out.

`drv` describes whether the car is front-wheel drive, rear wheel drive, or 4wd.

4. Make a scatterplot of `hwy` vs `cyl`.

```{r}
ggplot(mpg, aes(x = hwy, y = cyl)) +
  geom_point()
```

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r}
ggplot(mpg, aes(x = class, y = drv)) +
  geom_point()
```

Scatterplots are suitable for displaying continuous variables (e.g., `cty` and `hwy`). `class` and `drv` are discrete variables.

## 3.3.1 Exercises

1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

In this example, colour has been mapped to an aesthetic. However, "blue" is not a variable in the dataset. To make the points blue, colour must be set manually (i.e., located *outside* the aesthetic mappings):


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

3. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg?`

Categorical: `manufacturer`, `model`, `trans`, `drv`, `fl`, `class`  
Continuous: `year`, `cty`, `hwy`

You can see this information located underneath the variable name in the output (e.g., `<chr>`)

```{r}
mpg
```
