---
title: "Chapter 28: Graphics for communication"
output: html_document
---

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```

### 28.2.1 Exercises

1. Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.

```{r}
ggplot(mpg, aes(x = hwy, y = cty)) +
  geom_point(aes(colour = class)) +
  labs(
    title = "Small cars are more fuel efficient",
    subtitle = "SUVs have poor fuel economy",
    caption = "Data from fueleconomy.gov",
    x = "Highway miles per gallon",
    y = "City miles per gallon",
    colour = "Class"
  )
```

2. The geom_smooth() is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

3. Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

```{r}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(aes(colour = Species)) +
  labs(
    title = "Iris virginica and versicolor have large petals",
    subtitle = "Iris setosa have substantially smaller petals",
    caption = "Source: Edgar Anderson's Iris Data",
    x = "Petal length (cm)",
    y = "Petal width (cm)",
    colour = "Species of iris"
  )
```

### 28.3.1 Exercises

1. Use geom_text() with infinite positions to place text at the four corners of the plot.

