---
title: "11. Data import"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)
```

### 11.2.2 Exercises

1.  What function would you use to read a file where fields were separated with  
    "|"?

`read_delim()`. The `delim` argument can be used to seperate fields with "|".

2.  Apart from `file`, `skip`, and `comment`, what other arguments do
    `read_csv()` and `read_tsv()` have in common?

Some other arguments they have in common include `col_names`, `col_types`, `locale`, `na`. In fact, in appears they share the exact same arguments.

3.  What are the most important arguments to `read_fwf()`?

`widths`, `start`, and `end`.

4.  Sometimes strings in a CSV file contain commas. To prevent them from
    causing problems they need to be surrounded by a quoting character, like
    `"` or `'`. By convention, `read_csv()` assumes that the quoting
    character will be `"`, and if you want to change it you'll need to
    use `read_delim()` instead. What arguments do you need to specify
    to read the following text into a data frame?
    
    ```{r, eval = FALSE}
    "x,y\n1,'a,b'"
    ```

```{r}
x <- "x,y\n1,'a,b'"
df <- read_delim(x, ",", quote = "'")
```

5.  Identify what is wrong with each of the following inline CSV files. 
    What happens when you run the code?
    
    ```{r, eval = FALSE}
    read_csv("a,b\n1,2,3\n4,5,6")
    read_csv("a,b,c\n1,2\n1,2,3,4")
    read_csv("a,b\n\"1")
    read_csv("a,b\n1,2\na,b")
    read_csv("a;b\n1;3")
    ```

```{r}
read_csv("a,b,c\n1,2,3\n4,5,6")
```

```{r}
read_csv("a,b,c\n1,2,1\n2,3,4")
```

```{r, warning=FALSE}
read_csv("a,b\n1")
```

It's unclear what the intent was here:
```{r}
read_csv("a,b\n1,2\na,b")
```

```{r}
read_csv("a,b\n1,3")
```

### 11.3.5 Exercises

1.  What are the most important arguments to `locale()`?

`date_names` and maybe `decimal_mark`.

2.  What happens if you try and set `decimal_mark` and `grouping_mark` 
    to the same character? What happens to the default value of 
    `grouping_mark` when you set `decimal_mark` to ","? What happens
    to the default value of `decimal_mark` when you set the `grouping_mark`
    to "."?

```{r, error=TRUE}
locale(decimal_mark = ",", grouping_mark = ",")
```

It produces an error.

```{r}
locale(decimal_mark = ",")
```

The default value for `grouping_mark` is changed to ".".

```{r}
locale(grouping_mark = ".")
```

The default value for `decimal_mark` is changed to ",".

3.  I didn't discuss the `date_format` and `time_format` options to
    `locale()`. What do they do? Construct an example that shows when 
    they might be useful.

`date_format` and `time_format` allows you to specify the default date and time formats.

```{r}
str(parse_guess("01/13/2018", locale = locale(date_format = "%m/%d/%Y")))
```

According to `vignette("readr")`, `time_format` isn't currently used for anything.

4.  If you live outside the US, create a new locale object that encapsulates 
    the settings for the types of file you read most commonly.

```{r}
locale(date_names = "en", date_format = "%d/%m/%Y", decimal_mark = ".",
       grouping_mark = ",", tz = "Australia/Perth")
```

5.  What's the difference between `read_csv()` and `read_csv2()`?

`read_csv2` uses ";" for separators instead of ",".

6.  What are the most common encodings used in Europe? What are the
    most common encodings used in Asia? Do some googling to find out.

UTF-8 is the standard worldwide. In Europe, ISO 8859-1 is a common encoding. Shift JIS is a common encoding in Asia.

7.  Generate the correct format string to parse each of the following 
    dates and times:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

```{r}
str(parse_date(d1, format = "%B %d, %Y"))
str(parse_date(d2, format = "%Y-%b-%d"))
str(parse_date(d3, format = "%d-%b-%Y"))
str(parse_date(d4, format = "%B %d (%Y)"))
str(parse_date(d5, format = "%m/%d/%y"))
str(parse_time(t1, format = "%H%M"))
str(parse_time(t2, format = "%H:%M:%OS %p"))
```