---
title: "5. Data transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r, include=FALSE}
library(nycflights13)
library(tidyverse)
``` 

### 5.2.4 Exercises

1.  Find all flights that  

    1. Had an arrival delay of two or more hours
    2. Flew to Houston (`IAH` or `HOU`)
    3. Were operated by United, American, or Delta
    4. Departed in summer (July, August, and September)
    5. Arrived more than two hours late, but didn't leave late
    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    7. Departed between midnight and 6am (inclusive)

Had an arrival delay of two or more hours:
```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (`IAH` or `HOU`):
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

Were operated by United, American, or Delta:

```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```
(To find the two letter carrier abbreviations, see `airlines`.)

Departed in summer (July, August, and September):
```{r}
filter(flights, month %in% c(7, 8, 9))
```

Arrived more than two hours late, but didn't leave late:
```{r}
filter(flights, arr_delay >= 120 & dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in flight:
```{r}
filter(flights, dep_delay >= 60  & dep_delay - arr_delay > 30)
```

Departed between midnight and 6am (inclusive):
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

2.  Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

It is a shortcut for finding values that fall within a specified range:

```{r}
filter(flights, between(month, 7, 9))
```

3.  How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

8,255 flights. 

To see what other variables are missing, you can use `View()`:

```{r, eval=FALSE}
View(filter(flights, is.na(dep_time)))
```

`dep_delay`, `arr_time`, `arr_delay`, and `air_time` are also missing. These flights were most likely cancelled.

4.  Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

See Jeffrey Arnold's [explanation](https://jrnold.github.io/r4ds-exercise-solutions/data-transformation.html#filter).

### 5.3.1 Exercises

1.  How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
arrange(flights, desc(is.na(dep_time)))
```

2.  Sort `flights` to find the most delayed flights. Find the flights that left earliest.

```{r}
arrange(flights, desc(dep_delay))
```

```{r}
arrange(flights, dep_delay)
```

3.  Sort `flights` to find the fastest flights.

This depends how you define "fastest flights". It could refer to the least amount of `air_time`:

```{r}
arrange(flights, air_time)
```

4.  Which flights travelled the longest? Which travelled the shortest?

If we assume this refers to distance travelled:

```{r}
arrange(flights, desc(distance))
```

The longest flight is from JFK to HNL (Honolulu) at 4983 miles.

```{r}
arrange(flights, distance)
```

The shortest flight is from EWR (Newark) to LGA (La Guardia) at 17 miles, however this appears to be a cancelled flight (a number of variables such as `dep_time` are missing). The next shortest flight is from EWR to PHL (Philadelphia) at 80 miles.

(See `airports` to find the names that correspond to the FAA airport code.)

### 5.4.1 Exercises

1.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, starts_with("dep"), starts_with("arr"))
```

2.  What happens if you include the name of a variable multiple times in a `select()` call?

It will only print once.

```{r}
select(flights, dep_time, dep_time)
```

3.  What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

The `one_of()` function allows you to use variables stored in a character vector:

```{r}
select(flights, one_of(vars))
```

It improves the readability of code.

4.  Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

Select helpers are not case sensitive by default. To change the default, use `ignore.case = FALSE`.

### 5.5.2 Exercises

1.  Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
transmute(flights, 
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100,
  dep_time_mins = (hour * 60) + minute
)

transmute(flights, 
  sched_dep_time,
  hour = sched_dep_time %/% 100,
  minute = sched_dep_time %% 100,
  sched_dep_mins = (hour * 60) + minute
)
```

2.  Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
transmute(flights, air_time, arr_time - dep_time)
```

You would expect `arr_time - dep_time == air_time`. This discrepancy is explained by timezone differences. `arr_time` and `dep_time` are in local timezones that may differ.

3.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

```{r}
select(flights, dep_time, sched_dep_time, dep_delay)
```

`dep_time - sched_dep_time == dep_delay`.

4.  Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for `min_rank()`.

```{r}
flights_ranked <- mutate(flights, rank = min_rank(-dep_delay))
arrange(flights_ranked, desc(dep_delay))
```

Ties should be assigned the lowest rank. `ties.method = "min"` is the default for `min_rank()`.

5.  What does `1:3 + 1:10` return? Why?

```{r, eval=TRUE}
1:3 + 1:10
```

1 + 1 = 2; 2 + 2 = 4; 3 + 3 = 6; 4 + 1 = 5; 5 + 2 = 7, and so on.

6. What trigonometric functions does R provide?

Cosine, sine, tangent, arc-cosine, and more. See `?Trig`.

### 5.6.7 Exercises

1.  Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:  

    + A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
  
    + A flight is always 10 minutes late.

    + A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

    + 99% of the time a flight is on time. 1% of the time it’s 2 hours late.  

Which is more important: arrival delay or departure delay?  

```{r}
flights %>% 
  group_by(flight, origin, dest) %>% 
  filter(arr_delay == 10) %>% 
  summarise()
```

```{r}
flights %>% 
  group_by(flight, origin, dest) %>%
  summarise(avg_arr_delay = mean(arr_delay, na.rm = TRUE), count = n()) %>% 
  filter(avg_arr_delay == 10)
```

Departure delay is probably more important for travellers.

2.  Come up with another approach that will give you the same output as `not_cancelled %>% count(dest)` and `not_cancelled %>% count(tailnum, wt = distance)` (without using `count()`).

```{r, include=FALSE}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(count = n())

not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(distance_total = sum(distance))
```

3.  Our definition of cancelled flights (`is.na(dep_delay) | is.na(arr_delay)`) is slightly suboptimal. Why? Which is the most important column?

Flights that do no depart cannot arrive at any location. However, flights that depart but are `NA` for `arr_delay` may have redirected or crashed.

The importance of the column depends on the question being answered. `dep_delay` is probably more representative of cancelled flights for the above reason.

4.  Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
flights %>% 
  group_by(month, day) %>% 
  summarise(cancelled = sum(is.na(dep_delay)))
```

It is difficult to tell just from this output. Let's visualise this instead:

```{r, include=TRUE}
flights %>% 
  mutate(cancelled = is.na(dep_delay)) %>% 
  group_by(month, day) %>% 
  summarise(
    prop_cancelled = mean(cancelled),
    avg_dep_delay = mean(dep_delay, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = prop_cancelled, y = avg_dep_delay)) +
    geom_point() +
    geom_smooth(se = TRUE)
```

There is a positive relationship between the proportion of cancelled flights and average delay. However, there are a few exceptional days with a high number of cancelled flights and relatively low average delay. If I had to guess, this might be related to bad weather (i.e., all flights were cancelled for a specific time period, so `avg_dep_delay` was close to zero).

5.  Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)

Frontier Airlines (F9) has the worst delays:

```{r}
not_cancelled %>% 
  group_by(carrier) %>% 
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(desc(avg_delay))
```

You cannot disentangle the effects of bad airports vs. bad carriers because there is an uneven distribution in destinations by carrier. Some carriers might operate more frequently from bad airports and vice versa.

6.  What does the `sort` argument to `count()` do. When might you use it?

`sort = TRUE` will sort output in descending order of `n`. You can use it for ranking instead of using `min_rank()` and/or `arrange()`.

```{r}
not_cancelled %>% count(dest, sort = TRUE)
```

### 5.7.1 Exercises

1.  Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

When combined with grouping, functions operate within groups only, not the entire dataset. For example:

```{r}
# What is the mean air time for each carrier?
flights %>%
  group_by(carrier) %>% 
  summarise(mean_air_time = mean(air_time, na.rm = TRUE))
```

```{r}
# What is the mean arrival delay for each month excluding all flights that were not delayed?
flights %>% 
  filter(arr_delay > 0) %>% 
  group_by(month) %>%
  summarise(avg_arr_delay = mean(arr_delay))
```

2.  Which plane (`tailnum`) has the worst on-time record?

N844MH.

```{r}
flights %>% 
  group_by(tailnum) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(avg_delay))
```

3.  What time of day should you fly if you want to avoid delays as much as possible?

Morning, preferably before 9 am.

```{r}
flights %>%
  group_by(hour) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(avg_delay)
```

4.  For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}
flights %>% 
  group_by(dest) %>% 
  summarise(total_delay = sum(arr_delay, na.rm = TRUE))
```

5.  Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using `lag()` explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  mutate(lag_delay = lag(dep_delay)) %>% 
  ggplot(aes(x = dep_delay, y = lag_delay)) +
    geom_point() +
    geom_smooth()
```

6.  Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

7.  Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}
flights %>% 
  group_by(dest, carrier) %>% 
  count(carrier) %>% 
  filter(carrier >= 2) %>% 
  group_by(carrier) %>% 
  count(sort = TRUE)
```

8.  For each plane, count the number of flights before the first delay of greater than 1 hour.

```{r}
flights %>% 
  group_by(flight) %>% 
  filter(dep_delay <= 60) %>% 
  count()
```